---
title: "STA 602 HW5"
author: "William Tirone"
format: pdf
editor: visual
---

```{r}
library(tidyverse)
```

# 4.1

Using beta - uniform conjugacy, we know the posterior will also be a beta distribution. So we have:

$$
p(\theta_2 | Y=30) \sim beta(30+1,20+1)\\
p(\theta_1 | Y=57) \sim beta(57+1,43+1)\\
$$

$$
P(\theta_1 < \theta_2 | data) = \frac{1}{S}\sum_{i=1}^{S}\bf{1} (\theta_1^{(S)} < \theta_2^{(S)})
$$

```{r}
theta1.sample = rbeta(5000,58,44)
theta2.sample = rbeta(5000,31,21)

mean(theta1.sample < theta2.sample)
```

# 4.2

## a)

The posteriors in 3.3 part a were given by:

$$
\theta_A | y_A \sim Gamma(237,20)\\
\theta_B|y_B \sim Gamma(125,14)
$$

Using the same methodology above with the indicator function:

```{r}
thetaA.sample = rgamma(5000,237,20)
thetaB.sample = rgamma(5000,125,14)

mean(thetaB.sample < thetaA.sample)
```

## b)

Our posterior for $\theta_B|y_B$ is $Gamma(\sum y + \alpha, \beta + n)$ so iterating through we will have $Gamma(117 + \alpha, \beta + 10)$. Observing the graph, it doesn't look like the event $\{\theta_A<\theta_B \}$ is incredibly sensitive, since it drops off fairly linearly over the iterations of n0. I would consider it sensitive if the probability dropped sharply at any n0, but it doesn't.

```{r}

# original data from 3.3
yB = c(11, 11, 10, 9, 9, 8, 7, 10, 6, 8, 8, 9, 7)
s = sum(yB)
n = length(yB)

# calculating the updates 
n0 = seq(1,75,1)
prob = c()

for (i in n0){
  thetaB.sample = rgamma(5000, s+(12*i),i+n)
  prob = c(prob,mean(thetaB.sample < thetaA.sample))
}

plot(n0,prob,col='#42c5f5',pch=17)
```
